### 2014-09-22 thesis/introduction/introduction.tex
Finished a preliminary version of the introduction #11

\section{Background}
Suppose that we have a (perhaps large) data set. It could consist of
demographic data taken from a census, details of transactions in a shopping
centre, or simply a collection of text documents. The term used to describe
the methods of analysis used to find unknown relationships and summarise
information in these data sets is \textit{data mining}. Broadly speaking,
data mining can be broken down into five areas (or \textit{tasks}) depending
on the aims of the person analysing the data:

@@

In this thesis we will focus on the task of predictive modelling. Specifically,
our problem will be a classification one. We will introduce the necessary
terminology in the next section as we describe the problem.

\subsection{Problem Statement}
Consider a data set of patients admitted to the trauma ward of a hospital.
This data set looks like a very large table: there are many columns, each one
describing some aspect of a patient (for example, age); there are also many
rows, each corresponding to a particular patient that was admitted. We will
refer to the columns as \textit{features} or \textit{attributes}, so-named
because they describe some particular aspect of the rows, which we will call
\textit{samples}. Concretely, as an example, our data set could contain
two features (age and gender) for five patients: as a table, this means that
we have two columns, one for age and another for gender, with five rows
containing the age and gender data for each patient. Attributes can be numeric
(such as age) and potentially take on an infinite number of different values,
or categorical (such as gender) and can only have one of a limited, pre-defined
list of values.

@@

\begin{enumerate}
\item The first problem is a \textit{binary} classification task in which we
predict one of two possible outcomes (or classes) for the patient: LOS less
than 2 days, or greater than 2 days. Given this, what model performs best?
\item The second problem is a \textit{multi-class} classification in which we
define several possible outcomes with different ranges for the LOS and predict
which range the patient will fall into. An example of possible outcomes can be
1-3 days, 4-10 days, and 11 or more days, giving us three classes for the
target variable. The optimal division of LOS into ranges will be investigated
as part of the problem.

\subsection{Previous Work}

\section{Contributions}

\section{Results and Evaluation}

>>>>

\section{Background}
Suppose that we have a (perhaps large) data set. It could consist of
demographic data taken from a census, details of transactions in a shopping
centre, or simply a collection of text documents. The term used to describe
the methods of analysis used to find unknown relationships and summarise
information in these data sets is \textit{data mining} \citep{Hand2001}.
Broadly speaking,
data mining can be broken down into five areas (or \textit{tasks}) depending
on the aims of the person analysing the data:

@@

In this thesis we will focus on the task of predictive modelling, also
commonly known as \textit{supervised learning}. Specifically,
our problem will be a classification one. We will introduce the necessary
terminology in the next section as we describe the problem.

\section{Problem Statement}
Consider a data set of patients admitted to the trauma ward of a hospital.
This data set looks like a very large table: there are many columns, each one
describing some aspect of a patient (for example, age); there are also many
rows, each corresponding to a particular patient that was admitted. We will
refer to the columns as \textit{features} or \textit{attributes}, so-named
because they describe some particular aspect of the rows, which we will call
\textit{samples}. Concretely, as an example, our data set could contain
two features (age and gender) for five patients: as a table, this means that
we have two columns, one for age and another for gender, with five rows
containing the age and gender data for each patient. Attributes can be numeric
(such as age) and potentially take on an infinite number of different values,
or categorical (such as gender) and can only have one of a limited, pre-defined
list of values.

@@

\begin{enumerate}
\item The first problem is a \textit{binary} classification task in which we
predict one of two possible outcomes (or classes) for the patient: LOS less
than 2 days, or greater than 2 days. Given this, what model performs best?
\item The second problem is a \textit{multi-class} classification in which we
define several possible outcomes with different ranges for the LOS and predict
which range the patient will fall into. An example of possible outcomes can be
1-3 days, 4-10 days, and 11 or more days, giving us three classes for the
target variable. The optimal division of LOS into ranges will be investigated
as part of the problem.
\end{enumerate}

The LOS is an important measure of resource utilisation in hospitals around
the world \citep{Ng2006},
and many studies have been carried out to investigate whether the
LOS of a patient in various medical domains -- after surgery, burns, or
intensive care treatment to name a few -- can be predicted accurately.
Regardless of the
specifics of the problem, accurate LOS prediction results in better planning
and management of limited resources such as staff, equipment and medical
supplies. This is particularly critical as hospitals often face the problem
of limited funding \citep{Walczak2003}.
As we will see in the next chapter, there has been limited
application of a wide range of data mining techniques to predict the LOS in
various medical domains, including trauma. Our work will aim to fill this
gap by providing insights on the application of data mining techniques to
LOS prediction.

\section{Contributions}
The work presented in this thesis aims to address several gaps in the current
research into LOS prediction, by building upon the work of Dinh et. al.
\citep{Dinh2013a}.
We apply and assess a range of techniques from different data mining paradigms
to a data set of trauma patients, and investigate whether formulating the LOS
prediction problem differently yields different results. The insights gained
from applying various data mining techniques will lead to a novel method of
predicting the LOS which will yield reliable and accurate results and support
the decision-making of medical professionals.

\section{Thesis Outline}
This thesis is structured as follows. Chapter \ref{chap:litreview} describes
the major methods in predictive modelling, interweaved with an analysis of
past work on LOS prediction. We explore in detail the major algorithms,
summarise the current state of research, and identify where our work fits
within this area.

Chapter \ref{chap:experiments} describes the experimental setup, detailing
each step in the process from data pre-processing to model creation and
finally to performance evaluation.

After the methods and setup are described, the results are presented and
discussed in Chapter \ref{chap:results}. We conclude in Chapter
\ref{chap:conclusion}, with indications for future work in this area.
